---
title: "Preliminary Analysis"
date: "`r Sys.Date()`"
output:
  bookdown::html_document2:
    theme: flatly
    code_folding: hide
    # toc: true
    # toc_depth: 2  
    # number_sections: true 
editor_options: 
  chunk_output_type: console
  markdown: 
    canonical: true
# bibliography: references.bib
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
ggplot2::theme_set(ggplot2::theme_minimal())

source(here::here("R", "environment.R"), echo = FALSE)
```

```{r load-data, include = FALSE}

all_trials <- read_csv(here("data", "reporting", "trials.csv"))
registrations <- read_csv(here("data", "reporting", "registrations.csv"))
all_results <- read_csv(here("data", "reporting", "results.csv"))
```

# Trial screening

This will need to be rerun after deduping registrations.

```{r screening}

# Returns list of inclusion counts and filtered dataframe
count_filter <- function(data, vars) {
  
  counts <- 
    tibble(name = as.character(),
           value = as.logical(),
           n = as.integer()
    )
  
  for (var in vars) {
    counts <-
      data %>% 
      count(.data[[var]]) %>% 
      pivot_longer(-n) %>% 
      add_row(counts, .)
    
    data <- filter(data, .data[[var]])
  }
  
  list(data = data, counts = counts)
  
}

# Screening criteria, in order
# Note: first 3 criteria alternatively encoded as `is_eligible_study`
screening_criteria <- c(
  "is_reg_2020",             # Must have 2020+ registration (auto)
  "is_intervention",         # Must be intervention/prevention (auto)
  "is_not_withdrawn",        # Must not be withdrawn (auto)
  "is_rcd_cutoff_1",         # Must be complete by cutoff (auto)
  "is_clinical_trial_manual",# Must be clinical trial (manual)
  "is_covid_manual",         # Must be covid intervention/prevention (manual)
  "is_not_withdrawn_manual", # Must not be withdrawn (manual)
  "is_not_dupe"              # Must not be duplicate
)

screened_trials <- 
  all_trials %>% 
  
  # is_dupe needs to be is_not_dupe
  # TODO: could change in back-end
  mutate(is_not_dupe = if_else(is_dupe, FALSE, TRUE, missing = TRUE)) %>% 
  count_filter(screening_criteria)

trials <- 
  screened_trials$data %>% 
  
  # Target enrollment should be numeric
  mutate(target_enrollment = as.numeric(target_enrollment))

screening_counts <- screened_trials$counts

kable(screening_counts)
```

The screened dataset for analysis includes `r nrow(trials)` trials.

# Results screening

*We may also want to report overall number of results, etc. but maybe not because we capture protocols, etc. and not clear what's what.*

For our analysis, we include limit results to:

-   only from trials in our screened dataset
-   only full-results and registry summary results
-   published by 2020-08-15 (excludes results without publication date) [Note: this excludes about half of our results]

```{r all-results}
# semi_join(trials, by = c("id_trial" = "id")) %>%
# filter(date_publication <= "2020-08-15") 

# Add results inclusion criteria
all_results <-
  
  all_results %>%
  
  mutate(
    is_trial_included = 
      if_else(id_trial %in% trials$id, TRUE, FALSE),
    
    is_pub_date_cutoff_1 = 
      if_else(date_publication <= "2020-08-15", TRUE, FALSE),
    
    is_pub_main =
      if_else(str_detect(pub_type, "other|interim", negate = TRUE), 
        TRUE, FALSE
      )
  )

results <-
  all_results %>% 
  filter(is_trial_included & is_pub_date_cutoff_1 & is_pub_main)
```

# Per Protocol Analysis

```{r completion-dates}

# This includes all publication completion dates (so could include from publications after 2020-08-15, interim results, etc.). We should decide whether to exclude.
# I checked. There are completion dates from 5 publications. All full results. 2 from after 2020-08-15.

# Prepare publication completion dates
# Remove the many publications without dates and spread dates for trials with multiple publications with dates (since some are the same and others are not)
pub_completion_dates <- 
  all_results %>% #or results? to exclude
  filter(!is.na(date_completion)) %>% 
  select(id_trial, date_completion) %>% 
  group_by(id_trial) %>% 
  mutate(pub_n = row_number()) %>% 
  pivot_wider(id_trial, 
              names_from = pub_n, names_prefix = "publication_",
              values_from = date_completion
  )
  
completion_dates <-
  trials %>% 
  select(id, auto = rcd_auto, manual = rcd_manual) %>% 
  left_join(
    pub_completion_dates,
    by = c("id" = "id_trial")
  ) %>% 

# TODO: decide what to report about completion dates
  mutate(
    # Does a publication report a completion date?
    has_pub_comp_date = if_else(!is.na(publication_1), TRUE, FALSE),
    
    # Does the completion date stay the same throughout?
    # If publications found, take into consideration.
    no_change = if_else(
    (auto == manual & is.na(publication_1) & is.na(publication_2)) |
      (auto == manual & manual == publication_1 & is.na(publication_2)) |
      (auto == manual & manual == publication_1 & publication_1 == publication_2),
      TRUE, FALSE
  ),
  
  # Did the registry date get postponed? Aka, is the manual date later than the the auto date?
  reg_date_change = 
    case_when(
    auto == manual ~ "No change",
    auto < manual ~ "Delay",
    auto > manual ~ "Early"
  )
  ) 

# completion_dates %>% 
#   rowwise() %>% 
#   mutate(latest = max(auto, manual, publication_1, publication_2))

n_date_change <-
  completion_dates %>% 
  filter(!no_change) %>% 
  nrow()

# Probably want to report as proportion of trials with publications
n_pub_comp_date <-
  completion_dates %>% 
  filter(has_pub_comp_date) %>% 
  nrow()

```

```{r latest-completion-dates}

# Get the latest completion date, indicating source based on auto > manual > publication
# Note: using latest completion date we will have some funkiness like reporting before completion, e.g. tri02934 (NCT04316377) which has a verified completion date in 2025

pub_comp_dates <- 
  results %>% # limit to results in analysis
  filter(!is.na(date_completion)) %>% 
  select(id_trial, date_completion) %>% 
  group_by(id_trial) %>% 
  # mutate(source = row_number(),
  #        source = glue::glue("publication_{source}")
  # ) %>% 
  mutate(source = "publication") %>% 
  ungroup() %>% 
  rename(id = id_trial)

latest_comp_dates <-
  trials %>% 
  select(id, auto = rcd_auto, manual = rcd_manual) %>% 
  pivot_longer(c(auto, manual), 
               names_to = "source", 
               values_to = "date_completion"
  ) %>% 
  bind_rows(pub_comp_dates) %>% 
  group_by(id) %>% 
  mutate(
    latest_date_completion = max(date_completion),
    source = factor(source, 
                    levels = c("auto", "manual", "publication"
                               # "publication_1", "publication_2"
                    )
                    )
  ) %>% 
  filter(date_completion == latest_date_completion) %>% 
  select(-date_completion) %>% 
  arrange(source, .by_group = TRUE) %>% 
  distinct(id, .keep_all = TRUE) %>% 
  ungroup()
```

```{r additional-descriptives}

n_ongoing_trials <-
  screening_counts %>% 
  filter(name == "is_rcd_cutoff_1" & value == FALSE) %>% 
  pull(n)

```

At each timepoint, we will report:

a.  The total number of clinical trials on COVID-19:

    1.  That are completed per the completion date: `r nrow(trials)` completed trials (based on completion date on registry on 2020-06-30)

    2.  That are ongoing per the completion date: `r n_ongoing_trials` ongoing trials (based on completion date on registry on 2020-06-30)

    3.  That were otherwise excluded from the analysis: See flowchart or we should decide if we want to specify other screening numbers in text.

    4.  For which the completion date on the registry was superseded by additional data (i.e., information from full results publication): Completion dates changed (in either the manual registry check or in a results publication) for `r n_date_change` trials. We found completion dates reported in results publications for `r n_pub_comp_date` trials. When completion dates did change on the registry between our automated and manual screenings, they were usually moved up:

    ```{r registry-completion-dates}
    completion_dates %>% 
      count(reg_date_change) %>% 
      kable()
    ```

    ```{r interim-results}

    interim_results_overview <-
      all_results %>%
      filter(str_detect(pub_type, "interim")) %>% 
      count(is_trial_included, is_pub_date_cutoff_1)

    n_interim_results_included_trials <-
      interim_results_overview %>% 
      filter(is_trial_included, is_pub_date_cutoff_1) %>% 
      pull(n)
    ```

    5.  For which we identified interim results: For trials included and by the publication cutoff date of 2020-08-15, we found `r n_interim_results_included_trials` interim result. We found additional interim results that are otherwise excluded (see table).

```{r table-interim-results}
kable(interim_results_overview)
```

1.  The total number of patients enrolled/anticipated to be enrolled in each clinical trial: Target enrollment ranged from `r min(trials$target_enrollment)` to `r     max(trials$target_enrollment)` with a mean of `r mean(trials$target_enrollment)` (IQR = `r IQR(trials$target_enrollment)`)

```{r results}

# Prepare result subsets
# results <-
#   all_results %>% 
#   filter(is_trial_included & is_pub_date_cutoff_1 & is_pub_main)

# results_missing_dates <-
#   all_results %>% 
#   filter(is_trial_included & is_pub_main & is.na(is_pub_date_cutoff_1)
#            # (is_pub_date_cutoff_1 | is.na(is_pub_date_cutoff_1))
#   )

results_plus_miss_dates <-
  all_results %>% 
  filter(is_trial_included & is_pub_main & 
           (is_pub_date_cutoff_1 | is.na(is_pub_date_cutoff_1))
  )

results_all_dates <-
  all_results %>% 
  filter(is_trial_included & is_pub_main)

results_oct_1 <-
  all_results %>% 
  filter(is_trial_included & is_pub_main & date_publication <= "2020-10-01")

n_results_oct <- nrow(results_oct_1)

n_trials_w_results_oct <- 
  results_oct_1 %>% 
  distinct(id_trial) %>% 
  nrow()

results_all_dates <-
  all_results %>% 
  filter(is_trial_included & is_pub_main)

# Prepare numbers to report
n_trials_w_results <- 
  results %>% 
  distinct(id_trial) %>% 
  nrow()

results_per_trial <-
  results %>% 
  count(id_trial, name = "n_results")

n_results_per_trial <-
  results_per_trial %>% 
  count(n_results, name = "n_trials")

n_1_results_per_trial <-
  results_per_trial %>% 
  filter(n_results == 1) %>% 
  nrow()

n_2_results_per_trial <-
  results_per_trial %>% 
  filter(n_results == 2) %>% 
  nrow()

n_3_results_per_trial <-
  results_per_trial %>% 
  filter(n_results == 3) %>% 
  nrow()

n_results_by_type <-
  results %>% 
  count(pub_type)

n_results <- nrow(results)

n_preprints <- 
  results %>% 
  filter(str_detect(pub_type, "article")) %>% 
  nrow()

n_articles <- 
  results %>% 
  filter(str_detect(pub_type, "preprint")) %>% 
  nrow()

n_summary <-
  results %>% 
  filter(str_detect(pub_type, "summary")) %>% 
  nrow()

n_trials <- nrow(trials)

n_results_no_date <- nrow(results_plus_miss_dates) - n_results

# kable(n_results_per_trial)
```

c.  The total number/proportion of clinical trials on COVID-19 with a results publication across all results publication types examined by this study (i.e., registry results, preprints, or journal articles) and descriptive statistics on publication time and type.

## Overall dissemination rate

`r scales::percent(n_trials_w_results/n_trials)` (`r n_trials_w_results`/`r n_trials`) of trials indicated as completed in the registry by and on 30 June 2020 disseminated a total of `r n_results` results across all publication types examined by this study (i.e., registry results, preprints, or journal articles) by 15 August 2020, six weeks after the cutoff date.

We found `r n_results_no_date` additional otherwise eligible result without a publication date, which is a registry summary results in a spreadsheet uploaded to Google Drive for ChiCTR2000031630. We do not consider this result in the analysis.

We started our results search in early October and can therefore report on dissemination through 1 October 2020. By 1 October 2020 `r n_results_oct` results were reported by `r n_trials_w_results_oct` trials, bringing the reporting rate up to `r scales::percent(n_trials_w_results_oct/n_trials)`. For further analyses, we consider results through 15 August only.

## Types and pathways of publication

The `r n_results` result publications include `r n_summary` (`r scales::percent(n_summary/n_results)`) registry results, `r n_preprints` (`r scales::percent(n_preprints/n_results)`) full-results preprints, and `r n_articles` (`r scales::percent(n_articles/n_results)`) journal articles. The majority (`r n_1_results_per_trial`, `r scales::percent(n_1_results_per_trial/n_trials_w_results)`) of trials reporting results did so through a single publication type. `r n_2_results_per_trial` (`r scales::percent(n_2_results_per_trial/n_trials_w_results)`) reported in 2 types, and only `r n_3_results_per_trial` (`r scales::percent(n_3_results_per_trial/n_trials_w_results)`) reported in all three types.

```{r plot-pubs, echo = FALSE}

# We can plot some other info with pubtype
# Registrations complication because of cross registration
# For now country, collapsing multinational
# Or maybe target_enrollment
# Or registration

trials_info <-
  trials %>% 
  semi_join(results, by = c("id" = "id_trial")) %>% 
  select(id, retrospective_registration, countries) %>% 
  # Collapse multinational (i.e., with comma)
  mutate(countries =  if_else(str_detect(countries, ","), "Multinational", countries))


plot_pubs <-
  results %>% 
  select(id_trial, pub_type) %>% 
  left_join(trials_info, by = c("id_trial" = "id")) %>% 
  mutate(pub_type = case_when(
    pub_type == "full_results_preprint" ~ "Full Preprint",
    pub_type == "full_results_journal_article" ~ "Full Journal Article",
    pub_type == "summary_results" ~ "Registry Results",
  )) %>% 
  # ggplot(aes(x = pub_type, fill = countries)) +
  ggplot(aes(x = pub_type, fill = retrospective_registration)) +
  geom_bar(position = "dodge") +
  facet_wrap(~countries) +
  # geom_label(aes(label = nrow(pub_type)), y = 20)+
  # xlab("All Publication Types") +
  # ggstatsplot::ggbarstats()
  scale_fill_viridis_d() +
  labs(
    title = "COVID-19 Clinical Trial Results Publications",
    subtitle = "By country and retrospective registration status"
  )
theme(
    axis.title.y = element_blank()
  ) +
  ggthemes::theme_fivethirtyeight()

plot_pubs_upset <-
  results %>% 
  select(id_trial, pub_type) %>%
  distinct() %>% # Remove multiple of single publication type
  group_by(id_trial) %>% 
  mutate(pub_type = case_when(
    pub_type == "full_results_preprint" ~ "Preprint",
    pub_type == "full_results_journal_article" ~ "Journal Article",
    pub_type == "summary_results" ~ "Registry Results"
  )) %>% 
  summarise(pub_types = list(pub_type)) %>% 
  ggplot(aes(x = pub_types)) +
  geom_bar() +
  geom_text(stat='count', aes(label = after_stat(count)), vjust = -.5) +
    ggupset::scale_x_upset(
    sets = c("Preprint", "Journal Article", "Registry Results")
  ) +
  scale_y_continuous(breaks = NULL, #lim = c(0, 3), 
                     name = "") +
  # xlab("Publication Types") +
  labs(
    title = "COVID-19 Clinical Trial Results",
    subtitle = "Intersection of Publication Types"
  ) +
  ggthemes::theme_fivethirtyeight() +
  theme_combmatrix(combmatrix.label.make_space = TRUE,
                   combmatrix.label.text = element_text(size=15)
  )

trials_pubtype <-
  results %>% 
  select(id_trial, pub_type) %>%
  distinct() %>% 
  mutate(pub_type = case_when(
    pub_type == "full_results_preprint" ~ "Preprint",
    pub_type == "full_results_journal_article" ~ "Journal Article",
    pub_type == "summary_results" ~ "Registry Results"
  ))
    
pub_sets <-
  split(trials_pubtype$id_trial, trials_pubtype$pub_type)

plot_pubs_venn <-
  ggVennDiagram::ggVennDiagram(pub_sets) +
  scale_fill_viridis_c() +
  guides(fill = "none" ) +
  labs(
    title = "COVID-19 Clinical Trial Results",
    subtitle = "Intersection of Publication Types"
  )

# ggVennDiagram::get_region_items(plot_pubs_venn)

plot_pubs
plot_pubs_upset
plot_pubs_venn
```

a.  A Kaplan-Meier plot detailing the current median time from (primary) completion date to any publication and to each results publication type.

b.  Median time from preprint publication to journal publication per the Kaplan-Meier method. In the case of very small samples, we may adjust the reporting method to better describe the available data.

```{r}
# let's check out time to pub for each results, ignoring multiple results per trial
# let's use the auto registry completion date for now

time_to_pub <-
  results %>% 
  select(id_trial, pub_type, date_publication) %>% 
  left_join(latest_comp_dates, by = c("id_trial" = "id")) %>% 
  mutate(time_to_pub = date_publication - latest_date_completion) %>% 
  
  # Remove outlier
  filter(id_trial != "tri02934") %>% 
  mutate(time_to_pub_int = as.numeric(time_to_pub)) %>% 
  mutate(pub_type = case_when(
    pub_type == "full_results_preprint" ~ "Full Preprint",
    pub_type == "full_results_journal_article" ~ "Full Journal Article",
    pub_type == "summary_results" ~ "Registry Results",
  ))
  

plot_time_dot <-
  ggplot(time_to_pub, aes(y = time_to_pub, x = factor(pub_type))) +
  geom_dotplot(binaxis = "y", stackdir = "center")

# plot_time_violin <-
#   ggplot(time_to_pub, 
#          aes(y = time_to_pub_int, 
#              x = pub_type, 
#              fill = pub_type)) +
#   see::geom_violindot(
#     fill_dots = "black", 
#     size_dots = 40,
#     scale = "count",
#     # trim = FALSE
#   ) +
#   scale_fill_viridis_d() +
#   theme_modern() +
#   labs(
#     title = "Time to publication for different publication types",
#     y = "Time in days between\ntrial completion date and publication date",
#     x = "Publication Type",
#     caption = "Completion date reflects latest completion date available for the trial on the registry or in the publication. Negative values reflect that some trials publish prior to official completion, or fail to update or report completion dates."
#   ) +
#   guides(fill = FALSE)

#ggbeeswarm

plot_time_dot
# plot_time_violin
```

```{r registrations}

```
