---
title: "Numbers for preprint"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE)
knitr::opts_chunk$set()
source(here::here("R", "environment.R"), echo = FALSE)
```

```{r load-data, include = FALSE}
registrations <- read_csv(here("data", "reporting", "registrations.csv"))
```

```{r trial-screening}
all_trials <- read_csv(here("data", "reporting", "trials.csv"))
source(here("R", "analysis", "screen-trials.R"))
```


```{r all-results}

all_results <- 
  
  read_csv(here("data", "reporting", "results.csv")) %>% 
  
  # Add results inclusion criteria
  mutate(
    is_trial_included = 
      if_else(id %in% trials$id, TRUE, FALSE),
    
    is_pub_date_cutoff_1 = 
      if_else(date_publication <= "2020-08-15", TRUE, FALSE),
    
    is_pub_main =
      if_else(str_detect(pub_type, "other|interim", negate = TRUE), 
        TRUE, FALSE
      )
  )

results <-
  all_results %>% 
  filter(is_trial_included & is_pub_date_cutoff_1 & is_pub_main)

n_full_results <- nrow(results)
n_trials_w_results <- n_distinct(results$id)
p_trials_w_results <- n_trials_w_results/ n_not_dupe
```

# Methods

## Search dates

```{r qualtrics-search-dates}

qualtrics_raw <-
  read_csv(here("data", "raw", "qualtrics-results-search.csv"), skip = 1)

qualtrics_clean <-
  qualtrics_raw %>%

  slice(-1) %>%

  # Clean up datatypes and colnames
  readr::type_convert() %>%
  janitor::clean_names()

qualtrics_clean %>%
  filter(finished & progress == 100) %>%
  summarize(start = min(start_date), end = max(end_date))

```

# Results

## Trial screening

As of 30 June 2020, the ICTRP COVID-19 database contained `r n_ictrp` registered trials. Following all automated and manual exclusions, our final analysis dataset included `r n_not_dupe` completed clinical trials. Details of all exclusions is available in Figure X. Brief descriptive details of the `r n_not_dupe` trials are included in Table X.

## Trials summary table
```{r trials-summary}

cross_registrations <-
  registrations %>% 
  semi_join(trials, by = "id") %>%
  group_by(id) %>% 
  mutate(
    n_reg = n(),
    cross_registry = if_else(n_reg == 1, registry, "Cross-registered")
  ) %>% 
  distinct(id, .keep_all = TRUE) %>% 
  select(id, cross_registry) #%>% 
  # right_join(trials, by = "id")

tbl_trials <-
  trials %>%
  
  left_join(cross_registrations, by = "id") %>% 
  
  mutate(
    has_result = if_else(id %in% unique(results$id), "With Results", "Without Results"),
    
    # Collapse multinational (i.e., with comma)
    countries = if_else(str_detect(countries, ","), "Multinational", countries),
  ) %>% 
    
    # Group countries with fewer than 4 trials (overall)
    # add_count() %>%
    group_by(countries) %>%
    mutate(
      n_countries = n(),
      countries = if_else(n_countries <= 4, "Countries w/ <5 trials", countries)
    ) %>%
  
  select(has_result, 
         target_enrollment, 
         study_type, 
         phase, 
         cross_registry, 
         countries
  ) %>%

  gtsummary::tbl_summary(
    by = has_result, 
    type = all_continuous() ~ "continuous2",
    statistic = all_continuous() ~ c("{N_nonmiss}",
                                     "{median} ({p25}, {p75})", 
                                     "{min}, {max}"),
    missing = "no",
    label = list(
      target_enrollment ~ "Target enrollment",
      study_type ~ "Study type", 
      phase ~ "Phase", 
      cross_registry ~ "Registry", 
      countries ~ "Countries"
    ),
    sort = list(
      cross_registry ~ "frequency", 
      countries ~ "frequency"
    ),
  ) %>%
  
  add_overall() %>% 
  
  add_difference(include = target_enrollment) %>%

  # modify_header(label = "**Publication Type**") %>%

  bold_labels() %>% 
  modify_spanning_header(all_stat_cols() ~ "**Results Disseminated**") %>% 
  modify_caption("**Trial Characteristics** (N = {N})") #%>% 
  #remove_row_type(study_type, type = "header")

# show_header_names(tbl)
tbl_trials
```

## Per protocol analysis



Using our per-protocol criteria, we located `r n_full_results` full results published before 16 August 2020 for `r n_trials_w_results`of the`r n_not_dupe`(`r style_percent(p_trials_w_results, symbol = TRUE)`) trials registered as completed by 30 June 2020. The breakdown of results by dissemination route is detailed in Figure X [The venn diagram]. Figure Y details the reporting of results by registry.

## Kaplan meier

Data for KM plots is available at [kaplan-meier-time-to-pub.csv](https://github.com/maia-sh/direcct/blob/master/data/reporting/kaplan-meier-time-to-pub.csv).

```{r kaplan-meier}
source(here("R", "analysis", "kaplan-meier.R"))
```

Figure Z shows time-to-publication using Kaplan-Meier (K-M) curves. Per protocol, we use the latest trial completion date reported in a publication when available (n = `r n_trials_pub_cd`), and otherwise the completion date reported in the registry, preferring the primary completion date to the study completion date, when available. For trials with multiple publications, we select the earliest publication date for the specified dissemination route. Trials with a publication date prior to the available completion date were considered reported at time 0 (n = `r n_any_pre` any publication; n = `r n_article_pre` journal articles, n = `r n_preprint_pre` preprints). We show time from completion to: a) first publication across any dissemination route, b) journal publication, and c) preprint publication. As no method of reporting crossed 50%, we could not generate a median time to publication. As only two trials reported results to a registry, we did generate a K-M plot for this route. These two trials were published `r n_days_summary` days after completion on the registry. For the `r n_preprint_article` trials that had both a preprint and a journal publication, the median days from first preprint publication to journal publication was `r median_preprint_article` (IQR = `r iqr_preprint_article`).

## Sensitivity Analyses
```{r sensitivity-analyses}
source(here("R", "analysis", "sensitivity-analyses.R"))
```
Whereas our per-protocol analysis relies on the trial completion date on the registry on 30 June 2020, we considered trial results reporting using the latest available completion date for the `r n_not_dupe` trials included in the main analysis. The 30 June 2020 registry completion date was superseded by a later completion date reported in a publication (regardless of whether the publication was included in our analysis) for n = `r n_latest_cd_pub` trials and a later completion date found on the registry during manual searches for n = `r n_latest_cd_manual` trials. Considering these updated completion dates, `r n_sa_trials_complete` trials were completed by 30 June per their latest available completion date, of which `r n_sa_trials_complete_results` trials (`r style_percent(p_sa_trials_complete_results, symbol = TRUE)`) had full results available by 16 August 2020.

Restricting our sample from the `r n_not_dupe` trials included in the main analysis to only those trials that reached full completion by 30 June 2020, showed `r n_sa_trials_scd_results` of `r n_sa_trials_scd` (`r style_percent(p_sa_trials_scd_results, symbol = TRUE)`) trials reported.

The inclusion of interim results adds an additional   `r n_sa_results_interim` result to our per-protocol analysis -- a reporting rate of `r style_percent(p_sa_trials_results_interim, symbol = TRUE)`.

Extending our reporting window beyond 6 weeks to 3 months (i.e., by 1 October 2020, prior to the start of our manual searches) adds an additional `r n_sa_results_3_mo - n_full_results` results for a reporting rate of `r n_sa_trials_results_3_mo`/`r n_not_dupe` (`r style_percent(p_sa_trials_results_3_mo, symbol = TRUE)`).

Data for 3-month follow-up KM plots is available at [kaplan-meier-time-to-pub.csv](https://github.com/maia-sh/direcct/blob/master/data/reporting/kaplan-meier-time-to-pub_3-mo.csv).
