---
title: "Preliminary Analysis"
date: "`r Sys.Date()`"
output:
  bookdown::html_document2:
    theme: flatly
    code_folding: hide
    # toc: true
    # toc_depth: 2  
    # number_sections: true 
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
# bibliography: references.bib
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
ggplot2::theme_set(ggplot2::theme_minimal())

source(here::here("R", "environment.R"))
```

```{r load-data, include = FALSE}

all_trials <- read_csv(here("data", "trials.csv"))
registrations <- read_csv(here("data", "registrations.csv"))
all_results <- read_csv(here("data", "results.csv"))
```

# Trial screening

This will need to be rerun after deduping registrations.

```{r screening}

# Returns list of inclusion counts and filtered dataframe
count_filter <- function(data, vars) {
  
  counts <- 
    tibble(name = as.character(),
           value = as.logical(),
           n = as.integer()
    )
  
  for (var in vars) {
    counts <-
      data %>% 
      count(.data[[var]]) %>% 
      pivot_longer(-n) %>% 
      add_row(counts, .)
    
    data <- filter(data, .data[[var]])
  }
  
  list(data = data, counts = counts)
  
}

# Screening criteria, in order
# Note: first 3 criteria alternatively encoded as `is_eligible_study`
screening_criteria <- c(
  "is_reg_2020",             # Must have 2020+ registration (auto)
  "is_intervention",         # Must be intervention/prevention (auto)
  "is_not_withdrawn",        # Must not be withdrawn (auto)
  "is_rcd_cutoff_1",         # Must be complete by cutoff (auto)
  "is_clinical_trial_manual",# Must be clinical trial (manual)
  "is_covid_manual",         # Must be covid intervention/prevention (manual)
  "is_not_withdrawn_manual"  # Must not be withdrawn (manual)
)

screened_trials <- 
  count_filter(all_trials, screening_criteria)

trials <- 
  screened_trials$data %>% 
  
  # Target enrollment should be numeric
  mutate(target_enrollment = as.numeric(target_enrollment))

screening_counts <- screened_trials$counts

kable(screening_counts)
```

The screened dataset for analysis includes `r nrow(trials)` trials.

# Results screening

*We may also want to report overall number of results, etc. but maybe
not because we capture protocols, etc. and not clear what's what.*

For our analysis, we include limit results to:

-   only from trials in our screened dataset
-   only full-results and registry summary results
-   published by 2020-08-15 (excludes results without publication date)
    [Note: this excludes about half of our results]

```{r all-results}
# semi_join(trials, by = c("id_trial" = "id")) %>%
# filter(date_publication <= "2020-08-15") 

# Add results inclusion criteria
all_results <-
  
  all_results %>%
  
  mutate(
    is_trial_included = 
      if_else(id_trial %in% trials$id, TRUE, FALSE),
    
    is_pub_date_cutoff_1 = 
      if_else(date_publication <= "2020-08-15", TRUE, FALSE),
    
    is_pub_main =
      if_else(str_detect(pub_type, "other|interim", negate = TRUE), 
        TRUE, FALSE
      )
  )

```

# Per Protocol Analysis

```{r completion-dates}

# This includes all publication completion dates (so could include from publications after 2020-08-15, interim results, etc.). We should decide whether to exclude.
# I checked. There are completion dates from 5 publications. All full results. 2 from after 2020-08-15.

# Prepare publication completion dates
# Remove the many publications without dates and spread dates for trials with multiple publications with dates (since some are the same and others are not)
pub_completion_dates <- 
  all_results %>% #or results? to exclude
  filter(!is.na(date_completion)) %>% 
  select(id_trial, date_completion) %>% 
  group_by(id_trial) %>% 
  mutate(pub_n = row_number()) %>% 
  pivot_wider(id_trial, 
              names_from = pub_n, names_prefix = "publication_",
              values_from = date_completion
  )
  
completion_dates <-
  trials %>% 
  select(id, auto = rcd_auto, manual = rcd_manual) %>% 
  left_join(
    pub_completion_dates,
    by = c("id" = "id_trial")
  ) %>% 

# TODO: decide what to report about completion dates
  mutate(
    # Does a publication report a completion date?
    has_pub_comp_date = if_else(!is.na(publication_1), TRUE, FALSE),
    
    # Does the completion date stay the same throughout?
    # If publications found, take into consideration.
    no_change = if_else(
    (auto == manual & is.na(publication_1) & is.na(publication_2)) |
      (auto == manual & manual == publication_1 & is.na(publication_2)) |
      (auto == manual & manual == publication_1 & publication_1 == publication_2),
      TRUE, FALSE
  ),
  
  # Did the registry date get postponed? Aka, is the manual date later than the the auto date?
  reg_date_change = 
    case_when(
    auto == manual ~ "No change",
    auto > manual ~ "Delay",
    auto < manual ~ "Early"
  )
  )

n_date_change <-
  completion_dates %>% 
  filter(!no_change) %>% 
  nrow()

# Probably want to report as proportion of trials with publications
n_pub_comp_date <-
  completion_dates %>% 
  filter(has_pub_comp_date) %>% 
  nrow()

```

```{r additional-descriptives}

n_ongoing_trials <-
  screening_counts %>% 
  filter(name == "is_rcd_cutoff_1" & value == FALSE) %>% 
  pull(n)

```

At each timepoint, we will report:

a.  The total number of clinical trials on COVID-19:

    1.  That are completed per the completion date: `r nrow(trials)`
        completed trials (based on completion date on registry on
        2020-06-30)

    2.  That are ongoing per the completion date: `r n_ongoing_trials`
        ongoing trials (based on completion date on registry on
        2020-06-30)

    3.  That were otherwise excluded from the analysis: See flowchart or
        we should decide if we want to specify other screening numbers
        in text.

    4.  For which the completion date on the registry was superseded by
        additional data (i.e., information from full results
        publication): Completion dates changed (in either the manual
        registry check or in a results publication) for
        `r n_date_change` trials. We found completion dates reported in
        results publications for `r n_pub_comp_date` trials. When
        completion dates did change on the registry between our
        automated and manual screenings, they were usually moved up:

    ```{r registry-completion-dates}
    completion_dates %>% 
      count(reg_date_change) %>% 
      kable()
    ```

    ```{r interim-results}

    interim_results_overview <-
      all_results %>%
      filter(str_detect(pub_type, "interim")) %>% 
      count(is_trial_included, is_pub_date_cutoff_1)

    n_interim_results_included_trials <-
      interim_results_overview %>% 
      filter(is_trial_included, is_pub_date_cutoff_1) %>% 
      pull(n)
    ```

    5.  For which we identified interim results: For trials included and
        by the publication cutoff date of 2020-08-15, we found
        `r n_interim_results_included_trials` interim result. We found
        additional interim results that are otherwise excluded (see
        table).

```{r table-interim-results}
kable(interim_results_overview)
```

1.  The total number of patients enrolled/anticipated to be enrolled in
    each clinical trial: Target enrollment ranged from
    `r min(trials$target_enrollment)` to
    `r     max(trials$target_enrollment)` with a mean of
    `r mean(trials$target_enrollment)` (IQR =
    `r IQR(trials$target_enrollment)`)

```{r results}

results <-
  all_results %>% 
  filter(is_trial_included & is_pub_date_cutoff_1 & is_pub_main)

n_trials_w_results <- 
  results %>% 
  distinct(id_trial) %>% 
  nrow()

n_results_per_trial <-
  results %>% 
  count(id_trial, name = "n_results") %>% 
  count(n_results, name = "n_trials")

```

c.  The total number/proportion of clinical trials on COVID-19 with a
    results publication across all results publication types examined by
    this study (i.e., registry results, preprints, or journal articles)
    and descriptive statistics on publication time and type.

    We found a total of `r nrow(results)` registry results, full-results
    preprints, and journal articles, for `r n_trials_w_results` trials.
    The majority of trials reporting results did so through a single
    publication type:

```{r n-results-per-trial}
kable(n_results_per_trial)
```

```{r plot-pubs}
plot_pubs <-
  results %>% 
  mutate(pub_type = case_when(
    pub_type == "full_results_preprint" ~ "Full Preprint",
    pub_type == "full_results_journal_article" ~ "Full Journal Article",
    pub_type == "summary_results" ~ "Registry Results",
  )) %>% 
  ggplot(aes(x = pub_type)) +
  geom_bar() +
  # xlab("All Publication Types") +
  ggtitle("COVID-19 Clinical Trial Results Publications") +
  theme(
    axis.title.y = element_blank()
  ) +
  ggthemes::theme_fivethirtyeight()

plot_pubs_combo <-
  results %>% 
  select(id_trial, pub_type) %>%
  distinct() %>% # Remove multiple of single publcation type
  group_by(id_trial) %>% 
  mutate(pub_type = case_when(
    pub_type == "full_results_preprint" ~ "Preprint",
    pub_type == "full_results_journal_article" ~ "Journal Article",
    pub_type == "summary_results" ~ "Registry Results"
  )) %>% 
  summarise(pub_types = list(pub_type)) %>% 
  ggplot(aes(x = pub_types)) +
  geom_bar() +
  geom_text(stat='count', aes(label = after_stat(count)), vjust = -.5) +
    ggupset::scale_x_upset(
    sets = c("Preprint", "Journal Article", "Registry Results")
  ) +
  scale_y_continuous(breaks = NULL, #lim = c(0, 3), 
                     name = "") +
  # xlab("Publication Types") +
  labs(
    title = "COVID-19 Clinical Trial Results",
    subtitle = "Intersection of Publication Types"
  ) +
  ggthemes::theme_fivethirtyeight() +
  theme_combmatrix(combmatrix.label.make_space = TRUE,
                   combmatrix.label.text = element_text(size=15)
  )

plot_pubs
plot_pubs_combo
```

a.  A Kaplan-Meier plot detailing the current median time from (primary)
    completion date to any publication and to each results publication
    type.

b.  Median time from preprint publication to journal publication per the
    Kaplan-Meier method. In the case of very small samples, we may
    adjust the reporting method to better describe the available data.
